### 2016-01-06 p2/abstract.tex
Fixed typos
New references and intro

With distributed computing and mobile applications,
synchronizing divergening replicas of data structures is a more and more common problem.
We use algebraic methods to reason about filesystem operations, 
and introduce a simplified definition of conflicting updates to filesystem trees.
We also define algorithms for update detection and reconciliation
and present rigorous proofs that they not only work as intended,
but also cannot be improved on.

@@

These results are often intuitive,
but providing exact proofs for them is far from trivial.
They contribute to our understanding of this special type of algebraic model,
and toward building a more complete algebra
of filesystems and other storage protocols.
They also form a theoretical basis for
specifying
and guaranteeing the error-free operation
of applications that implement an algebraic approach to synchronization.

>>>>

With distributed computing and mobile applications,
synchronizing diverging replicas of data structures is a more and more common problem.
We use algebraic methods to reason about filesystem operations, 
and introduce a simplified definition of conflicting updates to filesystem trees.
We also define algorithms for update detection and reconciliation
and present rigorous proofs that they not only work as intended,
but also cannot be improved on.

@@

These results are often intuitive,
but providing exact proofs for them is far from trivial.
They contribute to our understanding of this special type of algebraic model,
and toward building more complete algebras
of filesystems 
and extedning algebraic approaches to other data storage protocols.
They also form a theoretical basis for
specifying
and guaranteeing the error-free operation
of applications that implement an algebraic approach to synchronization.

### 2016-01-06 p2/comparisons.tex
Fixed typos
New references and intro

As a test of the proposed reconciliation algorithm, 
we compare our definition of conflicting
updates to how conflicts are defined by Balasubramaniam and Pierce
in their state-based approach implemented in the Unison synchronizer (\cite{BP}).
As we noted in \cite{NREC}, the update detector they describe provides a safe estimate of nodes
(paths) at which updates occurred.
It marks some nodes as \emph{dirty} in a way that we know that at non-dirty nodes
the filesystems (replicas) have not changed between their common original state and current state
(see Definition 3.1.1 in \cite{BP}).
The \emph{dirty} marks are also up-closed, that is, all ancestor nodes of a dirty node
are also dirty (Fact 3.1.3 in \cite{BP}).
And finally, in our notation, 
a conflict is detected between replicas $\FS_A$ and $\FS_B$ at node $n$
if $n$ is marked as dirty in both $\FS_A$ and $\FS_B$, and
$\FS_A(n)\neq\FS_B(n)$, and $n$ does not point to a directory in both replicas.
(In other words, 
$\FS_A(n)\not\in\setd \vee \FS_B(n)\not\in\setd$; see section 4.1 in \cite{BP}.
Let us not that there is an alternative approach to defining an algorithm for
the same synchronizer by Pierce and Vouillon in \cite{PV}.)

It can be easily seen that due to an edge case, 
not all conflicts detected based on the above definition
entails a conflict based on our system; that is, \cite{BP} describes a more
conservative policy.
We use the example we described in \cite{NREC},
where $\nn=\parent(n)$ and $\FS(\nn)$ is a directory and $\FS(n)$ is a file, and
the two replicas are derived in the following way:
\begin{gather*}
\FS_A = \cfba{n} \cc \cdba{\nn} \aFS \\
\FS_B = \cfba{n} \aFS.
\end{gather*}
Then, $\nn$ is dirty in both replicas:
in $\FS_A$ it was modified, and in $\FS_B$ one of its descendants was modified.
Moreover, $\FS_A(\nn)\neq\FS_B(\nn)$ and $\FS_A(\nn)=\empt\not\in\setd$,
and therefore a conflict is detected at $\nn$.
Our reconciliation algorithm detects no conflicts;
instead, it propagates $\cdba{\nn}$ to $\FS_B$, which we think is as expected
and desired.

>>>>

As a test of the proposed reconciliation algorithm, 
we compare our definition of conflicting
updates to how conflicts are defined by Balasubramaniam and Pierce
in their state-based approach implemented in the Unison synchronizer (\cite{BP}).
As we noted in \cite{NREC}, the update detector they describe provides a safe estimate of nodes
(paths) at which updates occurred.
It marks some nodes as \emph{dirty} in a way that we know that at non-dirty nodes
the filesystems (replicas) have not changed between their common original state and current state
(see Definition 3.1.1 in \cite{BP}).
The \emph{dirty} marks are also up-closed, that is, all ancestor nodes of a dirty node
are also dirty (Fact 3.1.3 in \cite{BP}).
And finally, in our notation, 
a conflict is detected between replicas $\FS_A$ and $\FS_B$ at node $n$
if $n$ is marked as dirty in both $\FS_A$ and $\FS_B$, and
$\FS_A(n)\neq\FS_B(n)$, and $n$ does not point to a directory in both replicas.
(In other words, 
$\FS_A(n)\not\in\setd \vee \FS_B(n)\not\in\setd$; see section 4.1 in \cite{BP}.
Let us note that there is an alternative approach to defining an algorithm for
the same synchronizer by Pierce and Vouillon in \cite{PV}.)

It can be easily seen that due to an edge case, 
not all conflicts detected based on the above definition
entails a conflict based on our system; that is, \cite{BP} describes a more
conservative policy.
We use the example we described in \cite{NREC},
where $\nn=\parent(n)$ and $\FS(\nn)$ is a directory and $\FS(n)$ is a file, and
the two replicas are derived in the following way:
\begin{gather*}
\FS_A = \cfba{n} \cc \cdba{\nn} \aFS \\
\FS_B = \cfba{n} \aFS.
\end{gather*}
Then, $\nn$ is dirty in both replicas:
in $\FS_A$ it was modified, and in $\FS_B$ one of its descendants was modified.
Moreover, $\FS_A(\nn)\neq\FS_B(\nn)$ and $\FS_A(\nn)=\empt\not\in\setd$,
and therefore a conflict is detected at $\nn$.
(This behaviour is preserved in the more recent Harmony synchronizer.
See ``delete/delete conflicts'' in \cite{PSG,FGKPS}.)
Our reconciliation algorithm detects no conflicts;
instead, it propagates $\cdba{\nn}$ to $\FS_B$, which we think is as expected
and desired.

### 2016-01-06 p2/intro.tex
New references and intro

In this paper we investigate a command-based approach
to synchronizing filesystem trees.
Presented with multiple copies (replicas) of the same filesystem that have been independently modified,
we regard the aim of the synchronizer to modify these replicas further
to make them as similar as possible.
The synchronization algorithm we describe
follows the two main steps described by Balasubramaniam and Pierce (\cite{BP}):
update detection, where we identify modifications that have been applied to the replicas
since they diverged,
and reconciliation, where we identify updates that can be propagated to other replicas and do so.
The remaining updates, which could not be propagated, are conflicting updates;
resolving these requires a separate system or human intervention.

@@

A central problem of command-based synchronizers 
during both update detection and reconciliation
is the ordering (scheduling) of commands.
If update detection is based on comparing the original state of the filesystem
to its new state, then we can easily collect a set of updates,
but we need to order these in a way that they could be applied
to the filesystems without causing an error 
(for example, a directory needs to be created before a file can be created under it).
Similarly, during reconciliation when we consider all the updates that
have been applied to the different replicas, we need to find a way of
merging these and creating a global ordering that we can apply
to the filesystems.
In fact, an insight in \cite{NREC} is that
if the ordering changes the effect of two commands (updates),
then under certain circumstances they are not compatible and will give rise to conflicts.
Accordingly, in this paper command-pairs and their properties
play a special role.
Other research also shows the essential nature of ordering:
in IceCube multiple orders are tested to find an acceptable one \cite{KRSD},
or in Bayou reconciling updates happens by redoing them in an acceptable order \cite{TTPDSH}.
Saito and Shapiro offer a survey of so-called optimistic replication algorithms in which
ordering (scheduling) plays a central role (\cite{SSH}).

In \cref{sec_def} we define a model of filesystems and a set of commands
that we will use to describe updates and modifications.
This is followed by investigating the properties and behaviors
of command pairs in \cref{section_axioms}.
\Cref{sec_update} describes update detection, and it is
in \cref{ordering} that we describe how commands can be ordered
and prove that under some simple conditions all valid orders
have the same effect.

>>>>

In this paper we investigate a command-based approach
to synchronizing filesystem trees.
Presented with multiple copies (replicas) of the same filesystem that have been independently modified,
we regard the aim of the synchronizer to modify these replicas further
to make them as similar as possible.
The synchronization algorithm we describe
follows the two main steps described by Balasubramaniam and Pierce (\cite{BP}):
update detection, where we identify modifications that have been applied to the replicas
since they diverged and represent these as sequences of commands,
and reconciliation, where we identify commands that can be propagated to other replicas and do so.
The remaining commands, which could not be propagated, represent conflicting updates;
resolving these requires a separate system or human intervention.

@@

A central problem of command-based synchronizers 
during both update detection and reconciliation
is the ordering (scheduling) of commands.
If update detection is based on comparing the original state of the filesystem
to its new state, then we can easily collect a set of updates,
but we need to order these in a way that they could be applied
to the filesystems without causing an error 
(for example, a directory needs to be created before a file can be created under it).
Similarly, during reconciliation when we consider all the commands that
have been applied to the different replicas, we need to find a way of
merging these and creating a global ordering that we can apply
to the filesystems.
In fact, an insight in \cite{NREC} is that
if the ordering changes the effect of two commands,
then under certain circumstances they are not compatible and will give rise to conflicts.
Accordingly, in this paper command-pairs and their properties
play a special role
as we aim to find valid orderings on filesystem trees.

In terms of the usual classification of synchronizers (\cite{TSR, PV, SSH}),
our approach to reconciliation is therefore operation-based as we investigate sets of commands,
while the update detection algorithm makes the full synchronizer 
similar to state-based synchronizers 
as its starting point is the states of the replicase themselves.
Moreover, our ordering algorithms only depend on the properties of commands 
such as commutativity or idempotence,
and therefore they can be classified as semantic schedulers.
This is in opposition to syntactic scheduling, which uses 
the origin and original temporal order of updates, 
and which may give rise to otherwise resolvable conflicts (\cite{SSH}).
Ordering plays a central role in other operation-based synchronizers.
We refer to a survey by Saito and Shapiro of so-called optimistic replication algorithms (\cite{SSH});
or, in particular,
to IceCube, where multiple orders are tested to find an acceptable one (\cite{KRSD}, see also \cite{MPV}),
or Bayou, where reconciling updates happens by redoing them in a globally determined order \cite{TTPDSH}.

We start by defining a model of filesystems and a set of commands
that we will use to describe updates and modifications in \cref{sec_def}.
This is followed by investigating the properties and behaviors
of command pairs in \cref{section_axioms}.
\Cref{sec_update} describes update detection, and it is
in \cref{ordering} that we describe how commands can be ordered
and prove that under some simple conditions all valid orders
have the same effect.

### 2016-01-06 p2/references.tex
New reference
New references and intro

% C 8
\bibitem{CJ}
Russ Cox and William Josephson,
``File synchronization with Vector Time Pairs,''
MIT CSAIL, Cambridge, MA, 
Tech. Rep., MIT-CSAIL-TR-2005-014, 2005.
% Accessed 22 November 2015.
% http://hdl.handle.net/1721.1/30527.

@@

@@

>>>>

% C 8
\bibitem{CJ}
Russ Cox and William Josephson,
``File synchronization with Vector Time Pairs,''
MIT CSAIL, Cambridge, MA, 
Tech. Rep., MIT-CSAIL-TR-2005-014, 2005.
Accessed 22 November 2015.
http://hdl.handle.net/1721.1/30527.

% F 18
\bititem{FGKPS}
% Foster, J Nathan and Greenwald, Michael B and Kirkegaard, Christian and Pierce, Benjamin C and Schmitt, Alan
J. Nathan Foster et al.,
``Schema-directed data synchronization,''
U of Pennsylvania,
Philadelphia, PA,
Technical Reports (CIS),
Paper 5,
2005.
Accessed 23 November 2015.\\
http://repository.upenn.edu/cis\undersc reports/5.

@@

% M 19
\bibitem{MPV}
Vidal Martins and Esther Pacitti and Patrick Valduriez,
``Distributed semantic reconciliation of replicated data,''
in \emph{Proc. of CDUR.}
Paris, France: CNAM,
2005,
pp. 48--53.

@@

% P 9
\bibitem{PSG}
Benjamin C. Pierce, Alan Schmitt, and Michael B. Greenwald, 
``Bringing Harmony to Optimism: An Experiment in Synchronizing Heterogeneous Tree-Structured Data,''
U of Pennsylvania,
Philadelphia, PA,
Technical Reports (CIS),
Paper 39,
2004.
Accessed 23 November 2015.\\
http://repository.upenn.edu/cis\undersc reports/39.
